### 2018-02-16 Writing/Thesis/content/chapter-evaluation.tex
more re-working

% Evaluation Measures used & computed
As rankings returned by this oracle represent the ideal solution, the predicted rankings are compared with them through the measures presented in chapter \ref{sec:fundamentals}. These include the Kendall rank correlation coefficient, loss and best three loss. For all three measures, it is computed if any advantage or disadvantage one solutions has is significant, by means of the Mann-Whitney U and its P-value. The regression based rankers pose a special case: internally, they predict a performance value for each classifier, which therefore can be compared to the actual performance values to show how well they are predicted. This is done by computing the root mean square error between the predicted and actual performance values. In summary, it is desirable for the rankers to come as close as possible to the correct ranking, which is reflected by a high rank correlation and a low loss and root mean square error.

% Baseline
Furthermore, we are interested in the question as to what degree knowing about the properties of a data set influences the quality of rankings. Therefore, a ranker that is agnostic of the meta features of a query data set, that is that will always return the same ranking for any data set, is used as a baseline. More specifically, a best algorithm strategy that iteratively determines a ranking by counting the number of data sets where an algorithm is the best choice, is implemented. That means the first algorithm in the best algorithm ranking is the classifier that performs best on most data sets, the second is the best on most data sets when only rankings excluding the first algorithm are considered, and so forth. This baseline is evaluated in the same way as the preference rankers; as it does not predict performance values, the root mean square error of predicted performance values cannot bet computed. Ideally, rankings returned by the preference and regression based rankers should then be statistically significant better than the baseline.

\section{Experimental Setup}
In the following paragraphs, it is briefly summarized under which conditions the experiments which led to the results discussed in the next section where observed. This includes which classifiers, data sets and meta features where chosen in the evaluation, what the specifications of the machines the evaluation was executed on where, how the evaluation was carried out, and which specific regression based and preference based implementations where used.

% Which classifiers 
The classifiers considered in the rankings were the learning algorithms implemented in WEKA that are fit for classification, excluding meta and ensemble methods. The full list of the 22 classifiers can be gathered from Table \ref{tab:bestAlgorithmRanking}. They where all used in their default configurations as specified by WEKA.

% Which data sets
Data sets for the analysis where gathered from OpenML. From all data sets considered as 'active' on OpenML, which yields 2050 data sets that have been approved with no severe issues found for them so far \cite{openMLGuide}, all data sets that comply with the constraints of the learning problem at hand where selected. Only data sets with a defined, nominal target feature that are in the .ARFF format which where not specifically designed for streaming analysis\footnotemark{} where considered, which resulted in a reduced selection of 812 instances. Since this amount of data sets hindered the evaluation considerably, large data sets with more than 1000 instances or 100 features where removed as well, leading to a final selection of 448 data sets. Even though evaluation was only carried out on the shorter list, both the list of 812 and 448 data sets are included in the supplementary material for completeness.

@@

% How base performance data generated
The performance data of the classifiers for the selected data sets was generated on a linux cluster, each node consisting of two Intel(R) Xeon(R) E5-2670 processors running at 2.6GHz, 16GB RAM with usage of up to 150 nodes. There was a timeout of eight hours set for the computation of a single performance value of a classifier on a data set. For each of the classifiers, the predictive accuracy (the percentage of correctly classified instances in the data set) was recorded on each data set by means of five times stratified Monte-Carlo cross validation with a 70/30 split in training and test data. The stratified split was generated with the help of JAICore, 'A Java library for infrastructural classes and basic algorithms in the area reasoning, search, and machine learning' [Reference Missing]. When a timeout occurred or the evaluation failed otherwise, the predictive accuracy was set to zero.

% How rest of the evaluation carried out
All other experiments where carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) with 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts where used in this evaluation, and the rankers where evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values where recorded, they where trained with the meta data - performance examples for n - 1 data sets, and queried for the remaining data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML. Furthermore, the evaluation was carried out twice for all measures: once with the full meta data used, and one time using only the meta features not generated by landmarkers, that is the ones that require probing to be carried out for a new data set.

% Implementation of measures
For the computation of the Mann-Whitney U and Kendall rank correlation coefficient, Apache Commons implementations where used. Regarding the Mann-Whitney U, the strategy for ties as set to averaging, and the NaNs where set to be removed before the analysis was started.

% How find best version of preference and regression ranker
Furthermore, four alternatives where selected for a regression based and preference based alternative each, to get a general idea of how well the respective approach might be suited for the problem. For the regression based approach, the algorithms chosen where random forest, REPTree, M5P, and linear regression. For all of them, the WEKA implemenation \cite{hall2009weka} was used, and the algorithms where used with the standard hyperparameters set by WEKA. As the preference based ranking falls under the category of label ranking, and in the jPL-framework, which was used for the implementation of the label ranking algorithms and data set representations, only two alternatives for label ranking are implemented \cite{intelligent2017jpl}, the other two alternatives where generated by modifying the hyperparameters. The implemented alternatives are label ranking by pairwise comparison and instance based label ranking. As in early tests it became apparent that the instance based approach might hold more potential, for label ranking by pairwise comparison, only the default configuration dictated by jPL was used. For instance based label ranking, the default configuration, a configuration where the rank aggregation algorithm used by the label ranker was set to Kemeny-Young, and a configuration with Kemeny-Young rank aggregation and the number of neighbors of the base learner of the label ranker, namely k nearest neighbor (kNN), set to the square root of the number of instances in the training data set. In the following section, these eight variants are compared with the baseline and oracle mentioned in the previous section, and among each others.

@@

When looking at the results depicted in Table \ref{tab:evaluationResults}, it must first be noted that positive findings regarding the Kendall rank correlation can be reported. For the full set of meta data, six out of the eight implemented alternatives outperform the baseline, all of them significantly, as can be learned from Table \ref{tab:significanceResults} which shows the results of the significance tests, with random forest having the highest correlation of 0.495. In contrast, the two approaches that did not overtake the baseline, instance based label ranking in the default configuration and label ranking by pairwise comparison, show a significant disadvantage compared to the baseline, and thus also compared to the other ranking approaches. Regarding loss, only linear regression and random forest show a significant advantage, with all label ranking approaches except for instance based label ranking having a significant disadvantage in contrast. Moving on to best three loss, the picture largely remains the same, except that surprisingly, REPTree now also has a significant advantage over the baseline. The lowest value here is achieved by random forest with a best three loss of 1.308, in comparison to 2.440 for the best algorithm baseline. These values are illustrated in Figure \ref{fig:lossGraphics}, with a scatter plot for individual best three loss values on the data sets.

@@

Summarizing the observed results, in general, a regression based approach shows an indication to be stronger than a preference based approach. Specifically, using random forest as a regression model is the best solution observed in the evaluation, and label ranking by pairwise comparison the weakest, as it falls behind the other ranking approaches and the baseline often A surprising result is that the label ranking approaches achieve very similar results with or without meta data.

@@

Table \ref{tab:metaFeatureTimes} shows the times required for the computation of groups of meta features as discussed in chapter \ref{sec:approach}. As can be taken from the table, the computation of the full meta features takes approximately 135 milliseconds on average, while the simple meta features without probing add approximately 3 milliseconds to the prediction times on average. Regarding these computation times it is also notable that among the probing meta feature groups, some are considerably more expensive than others, so that in the future it might be sensible to consider only light probing with probing meta feature groups that are not as expensive.

@@

Further insights to be noted are the accuracy of predictions of performance values for classifiers by the regression models, and the static ranking constructed by the best algorithm baseline. The accuracy of the predictions achieved by the regression based rankers are depicted in Table \ref{tab:rootMeanSquareError}, for full and reduced meta data respectively. The first thing that attracts attention in this table is the absurdly high values for Linear Regression and M5P, however, these are all caused by the very high maximum value on one data set with id 685. If these extreme maxima remain out of consideration for Linear Regression and M5P (the other algorithms do not have this problem) more reasonable values are obtained. For M5P, on the full meta data the new maximum is 311.526, mean 8.127 and standard deviation 18.407; on the reduced meta data it is a maximum of 474.692, mean 15.444 and standard deviation 27.433. For Linear Regression, on the full meta data the new maximum is 142.996, mean 9.281 and standard deviation 8.438; on the reduced meta data it is a maximum of 131.456, mean 14.351 and standard deviation 10.299. However, these values are still worse than the ones obtained for the Random Forest and REPTree variants. It comes as little surprise that the Random Forest Ranker also delivers the most accurate results for performance estimates, but as these best values are more than 6\% off on average for the full meta data and even more than 10\% off on average for the reduced meta data, it is questionable whether these prediction can be useful. The usefulness is dependent on the task these predictions would be used for an whether if requires very accurate predictions of performance values.

@@

The second interesting thing to consider is the ranking which is created by the best algorithm baseline. This ranking constitutes meta-knowledge about the learning process itself in that it shows which algorithms are generally a good choice for many data sets. This ranking is depicted in Table \ref{tab:bestAlgorithmRanking}. The table is to be read top to bottom; of the 22 algorithms, the algorithm in the ith row is placed first on most data sets when compared to the next 22-i algorithms. The associated number represents the number of data sets on which the algorithm was placed first. It is, however, not a particularly surprising result, as for example the placement of Random Forest as the first algorithm was to be expected. Random Forest is a learning algorithm that is known to often deliver good performances, which was proven again in the results discussed previously - out of the reviewed ranking approaches, Random Forest was the best choice.

>>>>

% Evaluation Measures used & computed
As rankings returned by this oracle represent the ideal solution, the predicted rankings are compared with them through the measures presented in chapter \ref{sec:fundamentals}. These include the Kendall rank correlation coefficient, performance loss and best three loss regarding classifier performance. For all three measures, it is computed whether any difference in performance is significant, by means of the p-value for the Mann-Whitney U. The regression based rankers pose a special case: internally, they predict a performance value for each classifier, which therefore can be compared to the actual performance values to show how well they are predicted. This is done by computing the root mean square error between the predicted and actual performance values. In summary, it is desirable for the rankers to come as close as possible to the correct ranking, which is reflected by a high rank correlation and a low loss and root mean square error.

% Baseline
Furthermore, we are interested in the question as to what degree knowing about the properties of a data set influences the quality of rankings. Therefore, a ranker that is agnostic of the meta features of a query data set, that is that will always return the same ranking for any data set, is used as a baseline. More specifically, a best algorithm strategy that iteratively determines a ranking by counting the number of data sets where an algorithm is the best choice, is implemented. That means the first algorithm in the best algorithm ranking is the classifier that performs best on most data sets, the second is the best on most data sets when only rankings excluding the first algorithm are considered, and so forth. This baseline is evaluated in the same way as the preference rankers; as it does not predict performance values, the root mean square error of predicted performance values cannot bet computed. Ideally, rankings returned by the preference and regression based rankers should then be statistically significant better than the static baseline, as this would indicate a connection of certain meta features and a performance-induced ranking of classifiers for a data set.

\section{Experimental Setup}
In the following paragraphs, it is briefly summarized under which conditions the experiments which led to the results discussed in the next section were observed. This includes which classifiers and data sets were chosen, and what the specifications of the machines the evaluation was executed on were. It is described how the evaluation was carried out, and which specific regression-based and preference-based implementations where used.

% Which classifiers 
The classifiers considered in the rankings were the learning algorithms implemented in WEKA that are fit for classification, excluding meta and ensemble methods. The full list of the 22 classifiers can be gathered from Table \ref{tab:bestAlgorithmRanking}. They were all used in their default configurations as specified by WEKA.

% Which data sets
Data sets for the analysis were gathered from OpenML. From all data sets considered as 'active' on OpenML, which yields 2050 data sets that have been approved with no severe issues found for them so far \cite{openMLGuide}, all data sets that comply with the constraints of the learning problem at hand were selected. Only data sets in the .ARFF format with a defined, nominal target feature which were not specifically designed for streaming analysis\footnotemark{} were considered, which resulted in a reduced selection of 812 instances. Since this amount of data sets hindered the evaluation considerably, large data sets with more than 1000 instances or 100 features were removed as well, leading to a final selection of 448 data sets. Even though evaluation was only carried out on the shorter list, both the list of 812 and 448 data sets are included in the supplementary material for completeness.

@@

% How base performance data generated
The performance data of the classifiers for the selected data sets was generated on a linux cluster with nodes consisting of two Intel(R) Xeon(R) E5-2670 processors running at 2.6GHz, 64GB RAM. For the computation of a single performance value of a classifier on a data set, two CPU cores and 16GB RAM were used, with a timeout of eight hours. Up 150 of such processes were executed in parallel. For each of the classifiers, the predictive accuracy (the percentage of correctly classified instances in the data set) was recorded on each data set by means of five times stratified Monte-Carlo cross validation with a 70/30 split in training and test data. The stratified split was generated with the help of JAICore, 'A Java library for infrastructural classes and basic algorithms in the area reasoning, search, and machine learning' [Reference Missing]. When a timeout occurred or the evaluation failed otherwise, the predictive accuracy was set to zero.

% How rest of the evaluation carried out
All other experiments were carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) with 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts were used in this evaluation, and the rankers were evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values were recorded, they were trained with the meta data - performance examples for n - 1 data sets, and queried for the left-out data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML. Furthermore, the evaluation was carried out twice for all measures: once with the full meta data used, and one time using only the meta features not generated by landmarkers, that is the ones that require probing to be carried out for a new data set.

% Implementation of measures
For the computation of the Mann-Whitney U and Kendall rank correlation coefficient, Apache Commons implementations were used [REFERENCE MISSING]. Regarding the Mann-Whitney U, the strategy for ties as set to averaging, and the NaNs were set to be removed before the analysis was started.

% How find best version of preference and regression ranker
Furthermore, four alternatives were selected for a regression based and preference based alternative each, to get a general idea of how well the respective approach might be suited for the problem. For the regression based approach, the algorithms chosen were random forest, REPTree, M5P, and linear regression. For all of them, the WEKA implemenation \cite{hall2009weka} was used, and the algorithms where used with the standard hyperparameters set by WEKA. As the preference-based ranking falls under the category of label ranking, and in the jPL-framework, which was used for the implementation of the label ranking algorithms and data set representations, only two alternatives for label ranking are implemented \cite{intelligent2017jpl}, the other two alternatives were generated by modifying the hyperparameters. The implemented alternatives are label ranking by pairwise comparison and instance-based label ranking. As in early tests it became apparent that the instance-based approach might hold more potential, for label ranking by pairwise comparison, only the default configuration dictated by jPL was used. For instance-based label ranking, the default configuration, a configuration where the rank aggregation algorithm used by the label ranker was set to Kemeny-Young, and a configuration with Kemeny-Young rank aggregation and the number of neighbors of the base learner of the label ranker, namely k nearest neighbor (kNN), set to the square root of the number of instances in the training data set. All of the described rankers are included in an overview in Table \ref{tab:methods}. In the following section, these eight variants are compared with the baseline and oracle mentioned in the previous section, and among each others.

\begin{table}
\resizebox{\textwidth}{!}{%
\begin{tabularx}{1.25\textwidth}{>{\hsize=.4\hsize}X | >{\hsize=1.6\hsize}X}
	Ranker				& Configuration \\ \hline\hline
	LinearRegression		& Linear regression with S:0, R:1.0E-8, num-decimal-places:4 \\ \hline
	M5P					& M5 model trees with M:4 \\ \hline
	RandomForest			& Random forest with P:100, I:100, num-slots:1, K:0, M:1.0, V:0.001, S:1 \\ \hline
	REPTree				& REPT tree with M:2, V:0.001, N:3, S:1, L:1, I: 0.0 \\ \hline \hline	
	InstanceBased		& Instance-based label ranking with rankaggregation: Placket-Luce (norm tolerance: 1E-9, max iterations: 1000, log likelihood tolerance: 1E-5), baselearner: kNN (k: 10) \\ \hline
	InstanceBased		& Instance-based label ranking with rankaggregation: Kemeny-Young, baselearner: kNN (k: 10) \\ \hline
	InstanceBased		& Instance-based label ranking with rankaggregation: Kemeny-Young, baselearner: kNN (k:$\sqrt{\text{number of instances}}$) \\ \hline
	PairwiseComparison 	& Label ranking by pairwise comparison with baselearner: logistic regression (learning rate: 0.001), gradient step: adam (beta1: 0.99, beta2: 0.999, epsilon: 10E-8) \\
\end{tabularx}
}
\caption{An overview over the different base algorithms used in the evaluated rankers.}
\label{tab:methods}
\end{table}

@@

Table \ref{tab:evaluationResults} shows the results of the evaluation, with the best mean values and significant differences emphasized as described in the caption of the table. When looking at these results, it must first be noted that positive findings regarding the Kendall rank correlation can be reported. For the full set of meta data, six out of the eight implemented alternatives outperform the baseline, all of them significantly, as can be learned from Table \ref{tab:significanceResults} which shows the results of the significance tests, with random forest having the highest correlation of 0.495. In contrast, the two approaches that did not overtake the baseline, instance based label ranking in the default configuration and label ranking by pairwise comparison, show a significant disadvantage compared to the baseline, and thus also compared to the other ranking approaches. Regarding performance loss, only linear regression and random forest show a significant advantage, with all label ranking approaches except for instance based label ranking having a significant disadvantage in contrast. Moving on to best three loss, the picture largely remains the same, except that surprisingly, REPTree now also has a significant advantage over the baseline. The lowest value here is achieved by random forest with a best three loss of 1.308, in comparison to 2.440 for the best algorithm baseline. These values are illustrated in Figure \ref{fig:lossGraphics}, with a scatter plot for individual best three loss values on the data sets.

@@

Summarizing the observed results, in general, a regression based approach shows an indication to be stronger than a preference based approach. Specifically, using random forest as a regression model is the best solution observed in the evaluation, and label ranking by pairwise comparison the weakest, as it falls behind the other ranking approaches and the baseline often. A surprising result is that the label ranking approaches achieve very similar results with or without probing.

@@

Table \ref{tab:metaFeatureTimes} shows the times required for the computation of groups of meta features as discussed in Chapter \ref{sec:approach}. As can be taken from the table, the computation of the full meta features takes approximately 135 milliseconds on average, while the simple meta features without probing add approximately 3 milliseconds to the prediction times on average. Regarding these computation times it is also notable that among the probing meta feature groups, some are considerably more expensive than others, so that in the future it might be sensible to consider only light probing with probing meta feature groups that are not as expensive.

@@

Further insights to be noted are the accuracies predicted for classifiers by the regression models, and the static ranking constructed by the best algorithm baseline. The accuracy of the predictions achieved by the regression based rankers are depicted in Table \ref{tab:rootMeanSquareError}, for full and reduced meta data respectively. The first thing that attracts attention in this table is the absurdly high values for Linear Regression and M5P, however, these are all caused by the very high maximum value on one data set with id 685. If these extreme maxima remain out of consideration for Linear Regression and M5P (the other algorithms do not have this problem) more reasonable values are obtained. For M5P, on the full meta data the new maximum is 311.526, mean 8.127 and standard deviation 18.407; on the reduced meta data it is a maximum of 474.692, mean 15.444 and standard deviation 27.433. For Linear Regression, on the full meta data the new maximum is 142.996, mean 9.281 and standard deviation 8.438; on the reduced meta data it is a maximum of 131.456, mean 14.351 and standard deviation 10.299. However, these values are still worse than the ones obtained for the Random Forest and REPTree variants. It comes as little surprise that the Random Forest Ranker also delivers the most accurate results for performance estimates, but as these best values are more than 6\% off on average for the full meta data and even more than 10\% off on average for the reduced meta data, it is questionable whether these prediction can be useful. The usefulness is dependent on the task these predictions would be used for and whether it requires very accurate predictions of performance values.

@@

The second interesting thing to consider is the ranking which is created by the best algorithm baseline. This ranking constitutes meta-knowledge about the learning process itself in that it shows which algorithms are generally a good choice for many data sets. This ranking is depicted in Table \ref{tab:bestAlgorithmRanking}. The table is to be read top to bottom; of the 22 algorithms, the algorithm in the i-th row is placed first on most data sets when compared to the next 22-i algorithms. The associated number represents the number of data sets on which the algorithm was placed first. It is, however, not a particularly surprising result, as for example the placement of Random Forest as the first algorithm was to be expected. Random Forest is a learning algorithm that is known to often deliver good performances, which was proven again in the results discussed previously - out of the reviewed ranking approaches, Random Forest was the best choice.
